Integrating Environmental Forcing and Predator Effects into Antarctic Krill (*Euphausia superba*) Population Dynamics Using a Length-to-Age Stock Assessment Framework

### 2025-12-18 MS2_Krill.Rmd
improve intro

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
subtitle: "Version 2"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is fundamental for ecosystem-based fisheries management in the Western Antarctic Peninsula, where environmental variability and predator–prey interactions are known to strongly influence population productivity. However, most stock assessment frameworks applied to krill and other exploited marine resources are based on integrated models in which ecosystem processes are not explicitly embedded within the core population dynamics. In particular, environmental variability and predation pressure are seldom incorporated directly into the state equations governing recruitment, natural mortality, and biomass dynamics in integrated assessment models. Instead, these processes are typically explored through external sensitivity analyses or post hoc diagnostics, thereby limiting their influence on estimated stock status and management advice. In this study, an integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics. Within a Stock Synthesis–based framework, alternative model configurations are evaluated in which environmental covariates and predator indices are either explicitly linked to key state processes (recruitment and natural mortality) or excluded from the structural equations. This design allows a direct comparison of their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality. Systematic and coherent differences are observed among model configurations. When environmental variability and predator effects are embedded within the state equations, lower estimates of recruitment, total biomass, and spawning biomass are obtained, together with altered estimates of fishing mortality, relative to models in which ecosystem drivers are ignored. In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, less stable Monod-type parameter estimates, and more structured residual patterns. This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional ecological processes are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves. Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecological realism. Despite the challenges associated with integrating ecosystem information into fully integrated assessment models, it is demonstrated that environmental and predator effects can be consistently incorporated within the core population dynamics of krill assessments. The direct embedding of these drivers in the state equations provides a more precautionary and ecologically realistic representation of krill population dynamics and strengthens the scientific basis for management advice consistent with CCAMLR’s objectives for conservation and sustainable use in the Western Antarctic Peninsula.

@@

Antarctic krill (*Euphausia superba*, hereafter krill) is one of the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024]. Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Turner2005; @Lima2013; @Atkinson2019a; @Morley2020; @Carrasco2021; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales. It comprises three key components: (i) stock assessments to establish precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Warwick-Evans2022; @Kawaguchi2024].

@@

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations, which have traditionally been assessed from a data-limited context using model-based approaches. A wide variety of integrated models allow the use of diverse data sources, such as spatially explicit approaches including the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].

This study aims to integrate various data sources on krill (Euphausia superba), including fishery operations (e.g. catches and effort distribution), biological characteristics (e.g. growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g. environmental forcing and predator-prey interactions) into a unified stock assessment framework. Specifically, we use an integrated length-to-age model, which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. The primary objective is to evaluate the performance of alternative model configurations that represent different hypotheses about ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments. Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence for the value of ecosystem-informed models, showing that incorporating environmental and predator covariates has an impact on model fit, predictive skill, and structural uncertainty, and this has direct implications for sustainable management recommendations and precautionary krill harvesting strategies.

@@

One of the key challenges in this stock assessment framework is integrating ecosystem variables and drivers into the primary process of population dynamics [@Marshall2019] and SS3 addresses these approaches. According to [@methot2020stock], Stock Synthesis (SS3) can estimate recruitment levels as a function of the spawning biomass from the previous season, incorporating the effects of an environmental time series and a log-bias adjustment. This integrated approach allows for a more realistic and dynamic estimation of recruitment by linking it to both biological and environmental drivers in this way;

$$
E(\text{Recruitment}) = f(\text{SpBio}) \times \exp(B \times \text{envdata}) \times \exp\left(-0.5 \times \pi_R^2\right)
$$
$R$ represents the variability of deviations, adding to the variance caused by environmental factors. $SpBio$ represent spawning biomass, $envdata$ is a vector of time series of environmental variables. Consequently, as the environmental effect accounts for more of the total recruitment variability, the residual $R$ should be reduced. However, the model does not automatically adjust for this. Based on the results of linear models, we will identify the environmental variable that exhibits the strongest statistical influence on krill recruitment. This variable will be selected for inclusion in the stock assessment model to capture the environmental contribution to recruitment variability. By integrating the most influential environmental driver, the model is expected to enhance the biological realism of recruitment estimates and improve the accuracy of population dynamics estimates.

>>>>

---
title: "Integrating Environmental Forcing and Predator Effects into Antarctic Krill (*Euphausia superba*) Population Dynamics Using a Length-to-Age Stock Assessment Framework"
subtitle: "Version 2"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is fundamental for ecosystem-based fisheries management in the Western Antarctic Peninsula, where environmental variability and predator–prey interactions are known to strongly influence population productivity. However, most stock assessment frameworks applied to krill and other exploited marine resources are based on model-based approaches in which ecosystem drivers are not explicitly embedded within the core population dynamics. These drivers are typically explored through external sensitivity analyses or post-hoc diagnostics, thereby limiting their influence on estimated stock status and management advice. In this study, a spatial implicit integrated length-to-age assessment model is applied to Antarctic krill, allowing ecosystem drivers to be incorporated directly into the population dynamics. Within a Stock Synthesis–based framework, alternative model configurations are evaluated in which environmental covariates and predator indices are either explicitly linked to key state processes (recruitment and natural mortality) or excluded from the structural equations. This design allows assess their effects on recruitment, total biomass, spawning stock biomass, and fishing mortality. Systematic and coherent differences are observed among model configurations. When environmental variability and predator effects are embedded within the state equations, lower estimates of recruitment, total biomass, and spawning biomass are obtained, together with altered estimates of fishing mortality, relative to models in which ecosystem drivers are ignored. In addition, poorer statistical performance is generally observed for these ecosystem-informed models, as indicated by higher RMSE values, retrospective paterns and more structured residual patterns. This deterioration in fit is interpreted as the result of the increased informational and structural cost incurred by the model when additional components are imposed, rather than as evidence against the relevance of the ecological mechanisms themselves. Overall, these results indicate that stock assessments excluding ecosystem processes may overestimate population productivity and stock status by prioritizing statistical fit over ecosystem realism. Despite the challenges associated with integrating ecosystem information into fully integrated assessment models, it is demonstrated that environmental and predator effects can be consistently incorporated within the core population dynamics of krill assessments. The direct embedding of these drivers in the state equations provides a more precautionary and ecologically realistic representation of krill population dynamics and strengthens the scientific basis for management advice consistent with CCAMLR’s objectives for conservation and sustainable use in the Western Antarctic Peninsula.

@@

Antarctic krill (*Euphausia superba*, hereafter krill) is the largest populations on the planet in terms of biomass and also a keystone species in the Antarctic ecosystem [@Atkinson2019a; @McBride2021; @Kawaguchi2024]. Currently, the population is predominantly concentrated in the Western Antarctic Peninsula (WAP), being the primary area for krill harvesting over the past four decades and also an area that is also a critical feeding ground for predators such as penguins, seals, and whales [@Watters2013; @Hill2016; @Gallagher2023a]. At the same time, the WAP is one of the regions most affected by climate change, experiencing rapid increases in ocean temperature, sea ice loss, and changes in primary productivity [@Turner2005; @Lima2013; @Atkinson2019a; @Morley2020; @Carrasco2021; @McBride2021]. These overlapping pressures have intensified scientific interest in understanding krill population dynamics from both an ecological and fisheries perspective in such a complex and changing environment [@Warwick-Evans2022; @Kawaguchi2024].

Given the importance of krill to the ecosystem and its commercial value, effectively managing krill stocks is essential to maintaining the balance of the Antarctic ecosystem while supporting sustainable fisheries. Recognizing this, the Commission for the Conservation of Antarctic Marine Living Resources (CCAMLR) has been working on a revised ecosystem-based approach [@Constable2011; @Watters2013; @Cavanagh2016; @Constable2023; @Zhao2023; @Kawaguchi2024]. This strategy seeks to balance commercial harvesting with a precautionary framework that safeguards krill-dependent species in space and temporal scales. It comprises three key components: (i) stock assessments to establish precautionary harvest rates, (ii) periodic updates of krill biomass estimates at multiple spatial scales, and (iii) a spatial overlap analysis, to guide the spatial allocation of catches and mitigate ecosystem-wide risks [@Warwick-Evans2022; @Kawaguchi2024].

@@

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions, which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses in krill dynamic, and the interactions of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and their associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers, notably through CCAMLR’s Scheme of International Scientific Observation (SISO), which ensures standardized and independent biological sampling of the krill fishery across the Convention Area. Additionally, Multiple biomass assessment surveys based on hydroacoustic cruises are also conducted by CCAMLR, primarily led by countries such as the United States and China, whose programs represent the longest-running efforts and the broadest spatial coverage of krill hydroacoustic assessments in the Southern Ocean [@Hill2019; @Dornam2021; @Green2023]. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics. This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations.

CCAMLR has traditionally relied on assessments conducted within a data-limited context using relatively simple model-based approaches, such as GRYM-type models, which do not explicitly incorporate environmental or ecological variables in the representation of population dynamics. In contrast, a wide range of integrated assessment frameworks now allow the incorporation of diverse data sources and increasing levels of structural complexity. These include spatially explicit approaches such as the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. Notably, Stock Synthesis can also be applied to species that are difficult to age, such as Antarctic krill, through integrated length-to-age transformation models that explicitly link observed length compositions to underlying age-structured population dynamics. These advanced tools provide robust and flexible methodologies to address data limitations while accommodating complex biological processes and ecosystem interactions, thereby enabling a more realistic representation of population dynamics and improved scientific support for ecosystem-based fisheries management [@Hoyle2020].

This study has two main objectives. The first is to integrate multiple data sources on Antarctic krill (*Euphausia superba*), including fishery operations (e.g., catches and effort distribution), biological characteristics (e.g., growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g., environmental forcing and predator–prey interactions), into a unified stock assessment framework. Specifically, we apply an integrated length-to-age model that enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and overall productivity, as well as the potential influence of predator populations through the configuration of natural mortality. The second objective is to evaluate the performance of alternative model configurations representing different hypotheses regarding ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments. Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence of the value of ecosystem-informed assessment models, demonstrating that the incorporation of environmental and predator covariates influences model fit, predictive skill, and structural uncertainty, with direct implications for precautionary and sustainable management advice for the Antarctic krill fishery.

@@

To identify environmental drivers suitable for explicit inclusion in the stock assessment model, we evaluated the relationship between a krill recruitment index and three environmental variables known to influence krill population dynamics: Sea Surface Temperature (SST), Sea Ice Concentration (SIC), and Chlorophyll-a concentration (Chl-a) [@Flores2012; @Morley2020; @Saba2014; @Veytia2021; @Walsh2020]. These variables represent thermal, habitat, and trophic controls acting on early life stages of Antarctic krill.

Environmental data were obtained from a combination of satellite observations and reanalysis products (*Nimbus-7*, *ERA5*, and the EU Copernicus Marine Service). A summary of the spatial and temporal characteristics of each dataset is provided in Table @ref(tab:env). All variables were processed at their native spatial resolution using the R packages *sf* and *ncdf4* [@Pebezma2018; @Pierce2021], including data cleaning, spatial manipulation, rasterization, and coordinate transformation.

Prior to integration into the stock assessment model, environmental variables were statistically screened using a set of linear regression models with recruitment as the response variable. The baseline model accounted for spatial structure through management unit (MU), while subsequent models progressively incorporated Chl-a, SST, SIC, and selected interaction terms. This stepwise approach was used to identify the environmental variable exhibiting the strongest and most consistent statistical relationship with recruitment. Results of these screening analyses are provided in Supplementary Material 2.

### Environmental link formulation in Stock Synthesis

Stock Synthesis version 3.30.21 allows environmental covariates to be incorporated directly into the population dynamics through the *env-link* option in the control file, whereby external time series modify biological processes in a deterministic and multiplicative manner [@methot2020stock].

In this study, the selected environmental variable was linked to recruitment via an env-link acting on the recruitment deviations. Expected recruitment in year (t) was modeled as:

$$
E(R_t) = f(SSB_t) \times \exp(\beta_E E_t) \times \exp\left(-\tfrac{1}{2}\sigma_R^2\right),
$$

where ( f(SSB_t) ) represents the stock–recruitment relationship, assumed to follow a Beverton–Holt formulation; ( E_t ) corresponds to the standardized environmental time series provided in the Stock Synthesis data file; ( \beta_E ) denotes the environment–recruitment link coefficient estimated within the control file; and ( \sigma_R^2 ) is the variance associated with recruitment deviations.

On the log scale, this formulation is equivalent to:

$$
\log(R_t) = \log(f(SSB_t)) + \beta_E E_t + \eta_t,
\quad \eta_t \sim \mathcal{N}(0,\sigma_R^2).
$$

The environmental time series (E_t) was standardized prior to model fitting, such that (\beta_E) represents the change in log-recruitment associated with a one-standard-deviation change in the environmental variable.

The env-link coefficient (\beta_E) quantifies the sensitivity of recruitment to environmental variability:

$$
\beta_E = \frac{\partial \log(R_t)}{\partial E_t}.
$$

Positive values of (\beta_E) indicate environmental conditions that enhance recruitment, while negative values imply adverse effects. Because the env-link operates multiplicatively, environmental effects scale recruitment exponentially rather than additively.

By incorporating environmental forcing through the env-link, part of the interannual variability in recruitment that would otherwise be attributed to unexplained recruitment deviations is instead explained by the environmental covariate. Consequently, improved model fit is expected to manifest as a reduction in the estimated recruitment deviation variance (\sigma_R^2), although this reduction is not imposed a priori and arises through likelihood optimization.

The inclusion of the selected environmental variable via the env-link framework provides a biologically interpretable and statistically coherent mechanism to integrate ecosystem drivers into the core population dynamics of the krill stock assessment.
